Online Updating of Variational Bayes for Heterogeneous Forecasts of Vehicle Trajectory

### 2018-03-12 Cars/midcanPresentation.tex
presentation

\title[Online Updating of Variational Bayes]{Online Updating of Variational Bayes for Heterogeneous Forecasts of Vehicle Trajectory}
\author[Nathaniel Tomasetti]{Nathaniel Tomasetti}
\date{ }
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

\begin{frame}
\section{Background}
\subsection{Motivation}
\frametitle{Motivation}
\begin{itemize}
\item We aim to build a framework to provide density forecasts for new units in a heirarchy as they are observed.
\pause
\item Bayesian Inference in Forecasting:
\begin{enumerate}
\item Provides density forecasts for the variable of interest.
\item Provides a mechanism to update posterior distributions
\item Is easy to extend to hierarchical models.
\end{enumerate}
\pause
\vspace{3mm}
\item Bayesian Inference is only as good as its compuation:
\begin{enumerate}
\item Computation based on Markov Chain Monte Carlo (MCMC) is 'exact', but slow.
\item Computation based on Variational Bayes (VB) is fast, but approximate.
\end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Motivation}
\begin{itemize}
\item We focus on Hierarchical Models where:
\begin{enumerate}
\item Inference conditioned on existing units in the hierarchy is available.
\item Inference conditioned on additional units in a short time-frame is desired.
\end{enumerate}
\vspace{3mm}
\pause
\item We aim to infer, and forecast, additional units in an online data setting.
\pause
\item We want to leverage the existing information to improve the forecasts.
\end{itemize}
\end{frame}

\begin{frame}
\subsection{Hierarchical Models}
\frametitle{Hierarchical Models}
\begin{itemize}
\item Hierarchical Models are structured into three levels:
\begin{enumerate}
\item The likelihood: $p(z_i | \theta_i)$ for each unit $i$ in the model,
\item The prior: $p(\theta_i | \beta)$ where $\beta$ is shared across all $i$,
\item The hyper-prior: $p(\beta)$ is fixed.
\end{enumerate}
\pause
\vspace{5mm}
\item For $i = 1, 2, \dots, N$, we assume that the posterior distribution $$p(\beta, \theta_1, \dots, \theta_N | z_1, \dots, z_N)$$ is available.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Hierarchical Models}
\begin{itemize}
\item What about $i = N+1$? Using $\textbf{z}_{1:j} = \{z_i | i = 1, \dots, j\}$, $$p(\theta_{N+1} |\textbf{z}_{1:N+1}) \propto p(z_{N+1} | \theta_{N+1})\int_{\beta}p(\theta_{N+1} | \beta)p(\beta | \textbf{z}_{1:N})d\beta.$$
\pause
\item If there is a time constraint, we might use VB with the approximation $$q_{\lambda}(\theta_{N+1} | \textbf{z}_{1:N+1}) \approx p(\theta_{N+1} |\textbf{z}_{1:N+1})$$
\item choosing $$\lambda_{VB} = \arg \underset{\lambda}{\min}\mbox{ }D(p, q).$$
\end{itemize}
\end{frame}

\begin{frame}
\subsection{Variational Bayes}
\frametitle{Variational Bayes and the Kullback-Leibler Divergence}
\begin{itemize}
\item A common divergence function is the Kullback Leibler Divergence from $q$ to $p$,
\begin{align*}
\label{KL-def}
KL[q_{\lambda}(\theta_{N+1} &| \textbf{z}_{1:N+1})\hspace{.1cm}||\hspace{.1cm}p(\theta_{N+1} | \textbf{z}_{1:N+1})] = \\
&E_q \left[ \log(q_{\lambda}(\theta_{N+1} | \textbf{z}_{1:N+1})) - \log(p(\theta_{N+1} | \textbf{z}_{1:N+1})) \right],
\end{align*}
\pause
\item The expectation is intractable for most $p$ and $q$.
\item It typically cannot be estimated with Monte-Carlo simulations.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The Evidence Lower Bound}
\begin{itemize}
\item Instead look at the Evidence Lower Bound (ELBO), $$\mathcal{L}(q, \lambda) = E_{q} \left[\log(p(\theta_{N+1}, \textbf{z}_{1:N+1})) - \log(q_{\lambda}(\theta | \textbf{z}_{1:N+1}))\right],$$
\pause
\item Stochastic Estimates are available through Monte-Carlo: $$\mathcal{L}(q, \lambda) \approx \frac{1}{M} \sum_{j=1}^M \left(\log(p(\theta_{N+1}^{(j)}, \textbf{z}_{1:N+1})) - \log(q_{\lambda}(\theta_{N+1}^{(j)} | \textbf{z}_{1:N+1})) \right)$$ where $\theta_{N+1}^{(j)} \sim q_{\lambda}(\theta_{N+1} | \textbf{z}_{1:N+1})$.
\pause
\item Maximising the ELBO is equivalent to minimising the Kullback-Leibler Divergence.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Stochastic Optimisation}
\begin{itemize}
\item For an exponential family $p$ and factorisable $q$ Mean Field Variational Bayes provides a fast method to the ELBO. 
\item We do not restrict $p$ and $q$ this way and instead use Stochastic Gradient Ascent.
\item Repeat until $| \mathcal{L}(q, \lambda^{(m)}) - \mathcal{L}(q, \lambda{(m+1)})| < \nu$:
\begin{equation}
\label{gradientAscent}
\lambda^{(m+1)} = \lambda^{(m)} + \rho^{(m)} \widehat{\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda}} \bigg\rvert_{\lambda = \lambda^{(m)}}
\end{equation}
\end{itemize}
\begin{figure}
\centering
\includegraphics[width = 0.4\textwidth]{gradientascent}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Stochastic Optimisation}
\begin{itemize}
\item ELBO gradients can be estimated by the Score Estimator:
\begin{align*}
\widehat{\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda}}_{SC} &= \sum_{j = 1}^M \frac{\delta \log(q_{\lambda}(\theta_{N+1}^{(j)} | \textbf{z}_{1:N+1}))}{\delta \lambda}  \\
&\times \left(\log(p(\theta_{N+1}^{(j)}, \textbf{z}_{1:N+1})) - \log(q_{\lambda}(\theta_{N+1}^{(j)} | \textbf{z}_{1:N+1})) \right),
\end{align*}
where $\theta_{N+1}^{(j)} \sim q_{\lambda}(\theta_{N+1} | \textbf{z}_{1:N+1})$.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Stochastic Optimisation}
\begin{itemize}
\item Or the Reparameterised Estimator, where $\theta_{N+1} = f(\epsilon, \lambda)$:
\begin{align*}
\widehat{\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda}}_{RP} &= \sum_{j = 1}^M \frac{\delta J(\lambda, \epsilon^{(j)})}{\delta \lambda} + \frac{\delta \theta_{N+1}}{\delta \lambda} \\
\times \frac{\delta \log(p(\theta_{N+1}, \textbf{z}_{1:N+1}))}{\delta \theta_{N+1}} \bigg\rvert_{\theta_{N+1} = f(\lambda, \epsilon^{(j)})},
\end{align*}
where $\epsilon^{(j)} \sim q(\epsilon)$.
\end{itemize}
\end{frame}

\begin{frame}
\section{Online Heterogeneous Forecasting: Problem}
\subsection{Self-Driving Vehicles}
\frametitle{Self-Driving Vehicles}
\begin{itemize}
\item Self-Driving Vehicles are a thing soon
\item The navigation system has three steps
\begin{enumerate}
\item Detect surrounding traffic
\item Forecast surrounding traffic
\item Navigate through safely
\end{enumerate}
\item Forecasting has not had much work
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Forecasting Methods}
The literature forecasts vehicles in three ways:
\begin{enumerate}
\item Constant Behaviour.
\item Predict Events.
\item Provide Point Estimates.
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Aims of the first paper}
\begin{itemize}
\item What we plan on doing with this first paper
\item Overall goals and stuff
\item Motivate heterogeneity
\item Note heterogeneity needs inference on vehicles as they are encountered
\item There is not much time to do that
\item Develop an approximation to the Bayesian Update Equation so we can infer parameters for new vehicles online
\item Also mention something on density forecasting would be nice
\end{itemize}
\end{frame}

\begin{frame}
\subsection{The Next Generation Simulation Dataset}
\frametitle{Next Generation Simulation}
\begin{itemize}
\item This is a dataset created for macro traffic modelling
\item It records all of the cars all of the time
\item The road is curved
\item Before and After of road straightening project
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Trigonometric Motion Equations}
\begin{align}
x^*_{i, t} &= x^*_{i, t-1} + v_{i, t} \cos(\delta_{i, t}) \label{xEq}, \\
y^*_{i, t} &= y^*_{i, t-1} + v_{i, t} \sin(\delta_{i, t}) \label{yEq}.
\end{align}
Picture
\end{frame}

\begin{frame}
\frametitle{Extracting Driver Inputs}
\begin{align}
\delta_{i, t} &= 
     \begin{cases}
       \tan^{-1}\left(\frac{(y^*_{i, t} - y^*_{i, t-1})}{(x^*_{i, t} - x^*_{i, t-1})} \right)  &\quad\text{if }x^*_{i, t} \neq x^*_{i, t-1} \\
       \frac{\pi}{2} &\quad\text{if } y^*_{i, t} > y^*_{i, t-1} \mbox{ and } x^*_{i, t} = x^*_{i, t-1} \\
       -\frac{\pi}{2} &\quad\text{if } y^*_{i, t} < y^*_{i, t-1} \mbox{ and } x^*_{i, t} = x^*_{i, t-1} \\
       \delta_{i, t-1} &\quad\text{otherwise,} \\ 
     \end{cases} \label{dEq} \\
v_{i, t} &= \sqrt{(x^*_{i, t} - x^*_{i, t-1})^2 + (y^*_{i, t} - y^*_{i, t-1})^2} \label{vEq}.
\end{align}
\begin{equation}
\label{aEq}
a_{i, t} = v_{i, t} - v_{i, t-1}. 
\end{equation}
\begin{itemize}
\item Why other variables are useful
\item Acceleration is good because relative velocity / non-stationarity
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Forecasting in the framework}
\begin{itemize}
\item That is a terrible title.
\item Data is split into a training set and a test set:
\item Training vehicles are observed before driving,
\item Test vehicles are encountered while driving.
\end{itemize}
\begin{figure}
\centering
\includegraphics[width = \textwidth]{problem}
\end{figure}
\end{frame}

\begin{frame}
\subsection{Auto-regressive Models for Trajectory Forecasting}
\frametitle{Auto-regressive Models for Trajectory Forecasting}
\begin{figure}
\centering
\includegraphics[width = \textwidth]{pacf}
\end{figure}
\begin{itemize}
\item Dynamics are implemented in with bivariate AR models
\item There are three approaches:
\begin{enumerate}
\item The Homogeneous Model
\item The Independent Heterogeneous Model
\item The Clustered Heterogeneous Model
\end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{A Homogeneous Approach}
Talk about what is going on here
\begin{align}
a_{i, t} &= \sum_{j = 1}^p \phi_{j} a_{i, t-j} + \sigma_{\epsilon} \epsilon_{i, t} \label{aAR} \\
\delta_{i, t} &= \pi/2 + \sum_{j = 1}^q \gamma_{j} (\delta_{i, t-j} - \pi/2) + \sigma_{\eta} \eta_{i, t} \label{dAR}
\end{align}
\begin{equation*}
\label{thetaVec}
\theta = \{\phi_{1}, \dots, \phi_{p}, \gamma_{1}, \dots, \gamma_{q}, \log(\sigma^{2}_{\epsilon}), \log(\sigma^{2}_{\eta})\}.
\end{equation*}
\begin{equation}
\label{indPrior}
\theta \sim \mathcal{N}\left(\mu, \Sigma \right)
\end{equation}
\begin{itemize}
\item The Posterior
\item Why are we doing this
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The Independent Heterogeneity Model}
Maybe different drivers behave differently?
\begin{align}
a_{i, t} &= \sum_{j = 1}^p \phi_{i, j} a_{i, t-j} + \sigma_{\epsilon, i} \epsilon_{i, t} \label{aAR2} \\
\delta_{i, t} &= \pi/2 + \sum_{j = 1}^q \gamma_{i, j} (\delta_{i, t-j} - \pi/2) + \sigma_{\eta, i} \eta_{i, t}. \label{dAR2}
\end{align}
This breaks down into a bunch of independent posteriors
\begin{equation}
p(\theta_{1:N} \mid z_{1:N}) = \prod_{i=1}^N p(\theta_{i} \mid z_{i}),
\end{equation}
The training set isn't terribly useful.
\begin{equation}
p(\theta_{i}| \textbf{z}_{1:N}, z_{i,1:T}, \theta_{1:N}) = p(\theta_{i} | z_{i, 1:T}) \propto p(z_{i, 1:T} | \theta_i) p(\theta_i).
\label{indepNewCar}
\end{equation}
\end{frame}

\begin{frame}
\frametitle{The Clustered Heterogeneity Model}
Hierarchical Models share information 
\begin{itemize}
\item Introduce Auxiliary Variable K
\item Sharing of the information
\end{itemize}
\begin{equation}
\label{mixPrior}
\theta_i | k_i = j \sim N(\mu_j, \Sigma_j),
\end{equation}
and
\begin{equation}
k_i \mid \beta \sim \mbox{Multinomial}\left(\pi_1, \dots, \pi_{K}\right)
\end{equation}
\begin{align}
\mu_j &\sim \mathcal{N}\left(\bar{\mu}_j, \Omega_j\right), \\
\Sigma_j &\sim \mbox{Inverse Wishart}\left(\mbox{Degrees of Freedom } \tau_j, \mbox{Scale } \Psi_j\right), \\
\boldsymbol{\pi} &\sim \mbox{Dirichlet}\left(\alpha_1 = \alpha_2 = \dots = \alpha_K\right).
\end{align}
\end{frame}

\begin{frame}
\frametitle{The Clustered Heterogeneity Model}
The new vehicle depends on the old vehicles through $\beta$.
\begin{equation}
\label{hierNewCar}
p(\theta_{i} | \textbf{z}_{1:N}, z_{i, 1:T}) \propto p(z_{i, 1:T} | \theta_{i}) \int_{\beta} p(\theta_{i} | \beta) p (\beta | \textbf{z}_{1:N}) d\beta.
\end{equation}
The $\beta$ posterior is precise for large $N$
\begin{equation}
\label{hierNewCar2}
p(\theta_{i} | \textbf{z}_{1:N}, z_{i, 1:T}) \propto p(z_{i, 1:T} | \theta_{i}) p(\theta_{i} | \hat{\beta}).
\end{equation}
\end{frame}

\begin{frame}
\frametitle{Heterogeneity in the Model Posteriors}
\begin{figure}
\centering
\includegraphics[width = \textwidth]{HierSingleKDE}
\end{figure}
\end{frame}

\begin{frame}
\section{Updating Variational Bayes}
\frametitle{Updating Variational Bayes}
\begin{itemize}
\item Inference for self-driving vehicles needs to be fast.
\item Inference cannot slow down as the amount of data increases.
\begin{figure}[ht]
\centering
\includegraphics[width = \textwidth]{timeUpdate}
\end{figure}
\item How do we go from $q_{\lambda_S}(\theta_{i} | z_{i, 1:S})$ to $q_{\lambda_T}(\theta_{i} | z_{i, 1:T})$ for $S < T$?.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Approximating Bayesian Updating}
\begin{itemize}
\item The Bayesian Update Equation: 
\begin{equation}
\label{updatePost}
p(\theta_{i} | z_{i, 1:T}) \propto p(z_{i, S+1:T} | \theta_{i})p(\theta_{i} | z_{i, 1:S}).
\end{equation}
\item Computation requires evaluating the joint distribution,
\begin{equation}
\label{updateJoint}
p(\theta_{i}, z_{i, 1:T}) = p(z_{i, S+1:T} | \theta_{i})p(\theta_{i} | z_{i, 1:S}).
\end{equation}
\item We will approximate the Bayesian Update Equation in VB,
\begin{equation}
\label{ApproxJoint}
\hat{p}(\theta_{i},  z_{i, 1:T}) = p(z_{i, S+1:T} | \theta_{i})q_{\lambda_S}(\theta_{i} | z_{i, 1:S}).
\end{equation}
\item Source of additional error: How close $q$ is to $p$.
\end{itemize}
\end{frame}

>>>>

\title[Online Updating of Variational Bayes]{Online Updating of Variational Bayes for Heterogeneous Forecasts of Vehicle Trajectory}
\author[Nathaniel Tomasetti]{Nathaniel Tomasetti}
\date{ }
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

\AtBeginSection[]{
  \begin{frame}
  \vfill
  \centering
  \begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
    \usebeamerfont{title}\insertsectionhead\par%
  \end{beamercolorbox}
  \vfill
  \end{frame}
}

\begin{document}

@@

\section{Background}
\subsection{Motivation}

\begin{frame}
\frametitle{Motivation}
\begin{itemize}
\item We aim to build a framework to provide density forecasts for new units in a hierarchy as they are observed.
\pause
\item Bayesian Inference in Forecasting:
\begin{enumerate}
\item Provides density forecasts for the variable of interest.
\item Provides a mechanism to update posterior distributions
\item Is easy to extend to hierarchical models.
\end{enumerate}
\pause
\vspace{3mm}
\item Bayesian Inference is only as good as its compuation:
\begin{enumerate}
\item Computation based on Markov Chain Monte Carlo (MCMC) is 'exact', but slow.
\item Computation based on Variational Bayes (VB) is fast, but approximate.
\end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Motivation}
\begin{itemize}
\item We focus on Hierarchical Models where:
\begin{enumerate}
\item Each unit is a time series,
\item Inference conditioned on existing units in the hierarchy is available,
\item Inference conditioned on additional units in a short time-frame is desired.
\end{enumerate}
\vspace{3mm}
\pause
\item We aim to infer, and forecast, additional units as they are observed.
\pause
\item We want to leverage the existing information to improve the forecasts.
\end{itemize}
\end{frame}

\subsection{Bayesian Hierarchical Models}

\begin{frame}
\frametitle{Bayesian Hierarchical Models}
\begin{itemize}
\item Hierarchical Models are structured into three levels:
\begin{enumerate}
\item The likelihood: $p(z_i | \theta_i)$ for each unit $i$ in the model,
\item The prior: $p(\theta_i | \beta)$ where $\beta$ is shared across all $i$,
\item The hyper-prior: $p(\beta)$ is fixed.
\end{enumerate}
\pause
\vspace{5mm}
\item For $i = 1, 2, \dots, N$, we assume that the posterior distribution $$p(\beta, \theta_1, \dots, \theta_N | z_1, \dots, z_N)$$ is available.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Bayesian Hierarchical Models}
\begin{itemize}
\item What about $i = N+1$? Using $\textbf{z}_{1:j} = \{z_i | i = 1, \dots, j\}$, $$p(\theta_{N+1} |\textbf{z}_{1:N+1}) \propto p(z_{N+1} | \theta_{N+1})\int_{\beta}p(\theta_{N+1} | \beta)p(\beta | \textbf{z}_{1:N})d\beta.$$
\pause
\item If there is a time constraint, we might use VB with the approximation $$p(\theta_{N+1} |\textbf{z}_{1:N+1}) \approx q_{\lambda}(\theta_{N+1} | \textbf{z}_{1:N+1})$$
\item choosing $$\lambda_{VB} = \arg \underset{\lambda}{\min}\mbox{ }D(p, q).$$
\end{itemize}
\end{frame}

\subsection{Variational Bayes}

\begin{frame}[label=VB]
\frametitle{Variational Bayes}
\begin{itemize}
\item Typically minimize the Kullback-Leibler Divergence from $q$ to $p$,
\begin{align*}
\label{KL-def}
KL[q_{\lambda}(\theta_{N+1} &| \textbf{z}_{1:N+1})\hspace{.1cm}||\hspace{.1cm}p(\theta_{N+1} | \textbf{z}_{1:N+1})] = \\
&E_q \left[ \log(q_{\lambda}(\theta_{N+1} | \textbf{z}_{1:N+1})) - \log(p(\theta_{N+1} | \textbf{z}_{1:N+1})) \right],
\end{align*}
\item Maximising the Evidence Lower Bound (ELBO) is a tractable alternative, $$\mathcal{L}(q, \lambda) = E_{q} \left[\log(p(\theta_{N+1}, \textbf{z}_{1:N+1})) - \log(q_{\lambda}(\theta | \textbf{z}_{1:N+1}))\right],$$
\pause
\item For a fixed family $q$, find optimal $\lambda$ with \hyperlink{SGA}{gradient ascent}:
\begin{equation*}
\lambda^{(m+1)} = \lambda^{(m)} + \rho^{(m)} \widehat{\frac{\partial\mathcal{L}(q, \lambda)}{\partial \lambda}} \bigg\rvert_{\lambda = \lambda^{(m)}}
\end{equation*}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Stochastic Optimisation}
\begin{itemize}
\item ELBO gradients can be estimated by the Score Estimator:
\begin{align*}
\widehat{\frac{\partial\mathcal{L}(q, \lambda)}{\partial \lambda}}_{SC} &= \sum_{j = 1}^M \frac{\partial \log(q_{\lambda}(\theta_{N+1}^{(j)} | \textbf{z}_{1:N+1}))}{\partial \lambda}  \\
&\times \left(\log(p(\theta_{N+1}^{(j)}, \textbf{z}_{1:N+1})) - \log(q_{\lambda}(\theta_{N+1}^{(j)} | \textbf{z}_{1:N+1})) \right),
\end{align*}
where $\theta_{N+1}^{(j)} \sim q_{\lambda}(\theta_{N+1} | \textbf{z}_{1:N+1})$.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Stochastic Optimisation}
\begin{itemize}
\item Or the Reparameterised Estimator, where $\epsilon = f^{-1}(\theta_{N+1}, \lambda)$:
\begin{align*}
\widehat{\frac{\partial\mathcal{L}(q, \lambda)}{\partial \lambda}}_{RP} &= \sum_{j = 1}^M \frac{\partial J(\lambda, \epsilon^{(j)})}{\partial \lambda} + \frac{\partial \theta_{N+1}}{\partial \lambda} \\
&\times \frac{\partial \log(p(\theta_{N+1}, \textbf{z}_{1:N+1}))}{\partial \theta_{N+1}} \bigg\rvert_{\theta_{N+1} = f(\lambda, \epsilon^{(j)})},
\end{align*}
where $\epsilon^{(j)} \sim q(\epsilon)$ and $J(\lambda, \epsilon)$ is the Jacobian of $f$.
\end{itemize}
\end{frame}

\section{Online Heterogeneous Forecasting: Problem}
\subsection{Self-Driving Vehicles}

\begin{frame}
\frametitle{Self-Driving Vehicles}
\begin{itemize}
\item Many major manufacturers expect consumer self-driving vehicles to be available in the next 2-5 years
\pause
\item Autonomous driving has three major steps:
\begin{enumerate}
\item Detect surrounding traffic,
\item Forecast surrounding traffic,
\item Navigate through safely.
\end{enumerate}
\item We contribute to the second step.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Self-Driving Vehicles}
\begin{itemize}
\item The existing literature has three major categories:
\begin{enumerate}
\item \textbf{Constant Behaviour Models:} \cite{Gindele2010, Houenou2013, Bautista2017, Waymo2017}
\item \textbf{Predict Driver State:} \cite{Forbes1995, Pentland1999, Kuge2000, Gindele2010, Geng2017}
\item \textbf{Provide Point Estimates:} \cite{Wolf2008, Ding2013, Woo2016b, Geng2017, Woo2017}
\end{enumerate}
\pause
\item Our approach:
\begin{enumerate}
\item Provides density forecasts of vehicle trajectory,
\item Incorporates driver heterogeneity,
\item Provides feasible inference for online vehicle data.
\end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Forecasting Approach}
\begin{itemize}
\item \textbf{Training Vehicle} data is collected during development of the self-driving vehicle.
\item \textbf{Test Vehicles} are encountered while driving and require forecasts.
\begin{figure}
\centering
\includegraphics[width = \textwidth]{problem}
\end{figure}
\end{itemize}
\end{frame}

\subsection{The Next Generation Simulation Dataset}

\begin{frame}
\frametitle{Next Generation Simulation Data}
\begin{itemize}
\item The US Federal Highway Administration filmed 6101 vehicles every 100 milliseconds on the US 101 Freeway.
\item We model and remove the curvature from the dataset.
\begin{figure}
\centering
\includegraphics[width = \textwidth]{roadStraightening}
\end{figure}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Trigonometric Motion Equations}
\begin{itemize}
\item Changes in position are the result of velocity $v$ and steering angle $\delta$,
\begin{align}
x^*_{i, t} &= x^*_{i, t-1} + v_{i, t} \cos(\delta_{i, t}) \label{xEq}, \\
y^*_{i, t} &= y^*_{i, t-1} + v_{i, t} \sin(\delta_{i, t}) \label{yEq}.
\end{align}
\begin{figure}
\centering
\includegraphics[width = 0.7\textwidth]{motion}
\end{figure}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Extracting Driver Inputs}
\begin{itemize}
\item Transform the data to driver actions:
\end{itemize}
\begin{align}
\delta_{i, t} &= 
     \begin{cases}
       \tan^{-1}\left(\frac{(y^*_{i, t} - y^*_{i, t-1})}{(x^*_{i, t} - x^*_{i, t-1})} \right)  &\quad\text{if }x^*_{i, t} \neq x^*_{i, t-1} \\
       \frac{\pi}{2} &\quad\text{if } y^*_{i, t} > y^*_{i, t-1} \mbox{ and } x^*_{i, t} = x^*_{i, t-1} \\
       -\frac{\pi}{2} &\quad\text{if } y^*_{i, t} < y^*_{i, t-1} \mbox{ and } x^*_{i, t} = x^*_{i, t-1} \\
       \delta_{i, t-1} &\quad\text{otherwise,} \\ 
     \end{cases} \label{dEq} \\
v_{i, t} &= \sqrt{(x^*_{i, t} - x^*_{i, t-1})^2 + (y^*_{i, t} - y^*_{i, t-1})^2} \label{vEq}, \\
a_{i, t} &= v_{i, t} - v_{i, t-1}. \label{aEq}
\end{align}
\end{frame}

\begin{frame}
\frametitle{Vehicle Acceleration and Angle}
\begin{figure}
\includegraphics[width = \textwidth]{timeSeries}
\end{figure}
\end{frame}

\subsection{Auto-Regressive Models for Trajectory Forecasting}

\begin{frame}
\frametitle{Auto-Regressive Models for Trajectory Forecasting}
\begin{figure}
\centering
\includegraphics[width = \textwidth]{pacf2}
\end{figure}
\begin{itemize}
\item Dynamics are modelled with bivariate AR processes.
\pause
\item We consdier three approaches:
\begin{enumerate}
\item The Homogeneous Model
\item The Independent Heterogeneous Model
\item The Clustered Heterogeneous Model
\end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The Homogeneous Approach}
\begin{itemize}
\item Acceleration and Angle are modelled independently:
\begin{align}
a_{i, t} &= \sum_{j = 1}^p \phi_{j} a_{i, t-j} + \sigma_{\epsilon} \epsilon_{i, t}, \label{aAR} \\
\delta_{i, t} &= \pi/2 + \sum_{j = 1}^q \gamma_{j} (\delta_{i, t-j} - \pi/2) + \sigma_{\eta} \eta_{i, t}. \label{dAR} 
\end{align}
\pause
\item Collect parameters:
\begin{align}
\theta &= \{\phi_{1}, \dots, \phi_{p}, \gamma_{1}, \dots, \gamma_{q}, \log(\sigma^{2}_{\epsilon}), \log(\sigma^{2}_{\eta})\}. \label{thetaVec} \\
\theta &\sim \mathcal{N}\left(\mu, \Sigma \right) \label{indPrior} 
\end{align}
\item Parameters are shared across different vehicles.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The Independent Heterogeneity Model}
\begin{itemize}
\item Allow each vehicle to have an independent 
\item Each vehicle $i$ has an associated $\theta_i$:
\item Each $\theta_i$ has an identical, independent prior:
\begin{equation}
\theta_i \overset{iid}{\sim} \mathcal{N}\left(\mu, \Sigma \right). 
\end{equation}
\item Vehicle posteriors are independent.
\item The training set does not provide information about the test set.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The Clustered Heterogeneity Model}
\begin{itemize}
\item Replace the independent prior with a hierarchical prior:
\begin{equation}
\theta_i | \beta  \overset{iid}{\sim} \sum_{j=1}^K \pi_j \mathcal{N}(\mu_j, \Sigma_j)
\end{equation}
\pause
\item The hyper-prior is:
\begin{align*}
\mu_j &\sim \mathcal{N}\left(\bar{\mu}_j, \Omega_j\right), \\
\Sigma_j &\sim \mbox{Inverse Wishart}\left(\tau_j, \Psi_j\right), \\
\boldsymbol{\pi} &\sim \mbox{Dirichlet}\left(\alpha_1 = \alpha_2 = \dots = \alpha_K\right).
\end{align*}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The Clustered Heterogeneity Model}
\begin{itemize}
\item Test set vehicles depend on the train set vehicles through $\beta$.
\begin{equation}
\label{hierNewCar}
p(\theta_{N+1} | \textbf{z}_{1:N+1}) \propto p(z_{N+1} | \theta_{N+1}) \int_{\beta} p(\theta_{N+1} | \beta) p (\beta | \textbf{z}_{1:N}) d\beta.
\end{equation}
\pause
\item The $\beta$ posterior is precise for large $N$,
\begin{equation}
\label{hierNewCar2}
p(\theta_{N+1} | \textbf{z}_{1:N+1}) \propto p(z_{N+1} | \theta_{N+1}) p(\theta_{N+1} | \hat{\beta}).
\end{equation}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Evidence of Heterogeneity in the Model Posteriors}
\begin{figure}
\centering
\includegraphics[width = \textwidth]{hierSingleKDE2}
\end{figure}
\begin{itemize}
\item Top: Kernel Density Estimate of $E(\theta_i | z_i)$ from the IH Model.
\item Bottom: $p(\theta_{N+1} | \textbf{z}_{1:N})$ from the CH Model.
\end{itemize}
\end{frame}

\section{Updating Variational Bayes}

\begin{frame}
\frametitle{Updating Variational Bayes}
\begin{itemize}
\item Inference for self-driving vehicles needs to be fast.
\item Inference cannot slow down as the amount of data increases.
\begin{figure}[ht]
\centering
\includegraphics[width = \textwidth]{timeUpdate}
\end{figure}
\item Need to update $q_{\lambda_S}(\theta_{N+1} | z_{N+1, 1:S})$ to $q_{\lambda_T}(\theta_{N+1} | z_{N+1, 1:T})$.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Approximate Bayesian Updating}
\begin{itemize}
\item The Bayesian Update Equation: 
\begin{equation}
\label{updatePost}
p(\theta_{N+1} | z_{N+1, 1:T}) \propto p(z_{N+1, S+1:T} | \theta_{N+1})p(\theta_{N+1} | z_{N+1, 1:S}).
\end{equation}
\pause
\item Computation typically requires evaluating the joint distribution,
\begin{equation}
\label{updateJoint}
p(\theta_{N+1}, z_{N+1, 1:T}) = p(z_{N+1, S+1:T} | \theta_{N+1})p(\theta_{N+1} | z_{N+1, 1:S}).
\end{equation}
\pause
\item We use the VB approximation of the time $S$ posterior,
\begin{equation}
\label{ApproxJoint}
\hat{p}(\theta_{N+1},  z_{N+1, 1:T}) = p(z_{N+1, S+1:T} | \theta_{i})q_{\lambda_S}(\theta_{N+1} | z_{N+1, 1:S}).
\end{equation}
\item Source of additional error: How close $q$ is to $p$.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Updating Gradient Estimators}
\begin{itemize}
\item The updating score estimator:
\begin{align*}
&\widehat{\frac{\partial\mathcal{L}(q, \lambda_T)}{\partial \lambda_T}}_{USC} = \sum_{j = 1}^M \frac{\partial \log(q_{\lambda_T}(\theta_{N+1}^{(j)} | z_{N+1, 1:T}))}{\partial \lambda_T}  \\
&\times \left(\log(\hat{p}(\theta_{N+1},  z_{N+1, 1:T})) - \log(q_{\lambda_T}(\theta_{N+1}^{(j)} | z_{N+1, 1:T})) \right)
\end{align*}
where $\theta_{N+1}^{(j)} \sim q_{\lambda_T}(\theta_{N+1} | z_{N+1, 1:T})$. 
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Updating Gradient Estimators}
\begin{itemize}
\item The updating reparameterised estimator:
\begin{align*}
&\widehat{\frac{\partial\mathcal{L}(q, \lambda_T)}{\partial \lambda_T}}_{URP} = \sum_{j = 1}^M \frac{\partial J(\lambda_T, \epsilon^{(j)})}{\partial \lambda_T} + \frac{\partial f(\lambda_T, \epsilon^{(j)})}{\partial \lambda_T} \\
&\times \frac{\partial \log(\hat{p}(\theta_{N+1},  z_{N+1, 1:T})}{\partial \theta_{N+1}} \bigg\rvert_{\theta_{N+1} = f(\lambda_T, \epsilon^{(j)})} 
\end{align*}
where $\epsilon^{(j)} \sim q(\epsilon)$.
\end{itemize}
\end{frame}

\section{OHF: Results}
